### 2023-03-23 code/02-Ingest-R.qmd
add mess

---
title: "02-Ingest-R"
format: html
---

## Purpose

## Prompt

### Planning

3 parts:
- translation
- biographical information
- classification for our question (education, occupation)

#### Translation

Translate the following historical biography into English. It includes many Swedish abbreviations:

- Stored as string

### Biographical information

Convert this biography into schema.org/Person format.
* Return the output in JSON
* Include information on birthDate, birthPlace, education, occupation, family, parents. Return any other notable information in a field called "description" where possible.
* Include startDate and endDate wherever possible.

- Stored as JSON

### Classification for our question

Two further parts.

Education: Based on the information above, estimate the number of years of education that the individual received. Return your answer in JSON format with two fields, education_years (integer), education_justification (text).

Occupation: Based on the information above, Suggest 3 ISCO-08 occupational classifications and 3 O*NET occupational classifications. Return your answer in JSON format.

## Convert excel to YAML

```{r}
text <- readxl::read_excel(here::here("data", "test_50.xlsx"))

text <- text %>% 
  mutate(biography = str_squish(biography),
         id = row_number())
```

```{r}
library(yaml)

yaml <- text %>% as.yaml()

yaml %>% 
  write_yaml(here::here("data", "test_50.yaml"))
```

Test with 3

```{r}
yaml <- text %>% head(3) %>% as.yaml()

yaml %>% 
  write_yaml(here::here("data", "test_3.yaml"))
```

## Try with JSON instead

```{r}
text <- readxl::read_excel(here::here("data", "test_50.xlsx"))

text <- text %>% 
  mutate(biography = str_squish(biography),
         id = row_number())

json <- text %>%
  head(3) %>% 
  toJSON(pretty = TRUE)

json %>% 
  write_json(here::here("data", "test_3.json"))
```