### 2023-03-23 code/03-API-Interaction.qmd

---
title: "API Interaction"
format: html
---

## Purpose

Interact with the OPENAI API.

Step 1: translate the text.

```{r}
# Adapted from https://github.com/rasmusab/officialcocktails/blob/eb8f48ba2cd5c07f30bcd43262e967f4525a76de/01_regenerate-assets.R

library(tidyverse)
library(yaml)
library(jsonlite)
library(httr)
library(glue)

openai_api_key <- config::get("key")

```

```{r}
# Create the biographies_list if it doesn't already exist
# biographies <- read_yaml(here::here("data", "test_3.yaml"))

if(!file.exists(here::here("data", "biographies_list.yaml"))) {
  placeholder_biographies <- map(json, \(biography) {

    c(biography, list(
      translation = NULL
      # schema = NULL,
      # education = NULL,
      # occupation = NULL
      ))
  }) 
  write_yaml(placeholder_biographies, here::here("data", "biographies_list.yaml"))  
}

biographies_list <- read_yaml(here::here("data", "biographies_list.yaml"))
```

```{r}
# Calls the ChatGPT API with the given prompt and returns the answer
generate_text <- function(prompt) {
  response <- RETRY(
    "POST",
    times = 3,
    url = "https://api.openai.com/v1/chat/completions", 
    add_headers(Authorization = paste("Bearer", openai_api_key)),
    content_type_json(),
    encode = "json",
    body = list(
      model = "gpt-3.5-turbo",
      messages = list(list(
        role = "user", 
        content = prompt
      ))
    ),
    timeout(seconds = 30)
  )
  str_trim(content(response)$choices[[1]]$message$content)
}

# A function that takes a prompt template and generates another function which 
# takes a biography, as from biographies_list, completes the template and returns
# a ChatGPT answer.
gen_text_func <- function(...) {
  function(biography) {
    prompt <- glue(..., .envir = biography, .sep = "\n")
    cat(glue("Generating text for: {prompt}"), "\n")
    generate_text(prompt)
  }
}

# A list of generator functions that can generate text/images for the corresponding
# field, if missing.
generators <- list( 
  translation = gen_text_func("Translate the following historical biography into English. It includes many Swedish abbreviations: \n {biography}")
)

# Looping through the cocktail description and filling in all the blanks.
# As both the text and image generating APIs errors out now and then
# this loop saves every single response directly to cocktail-descriptions.yaml
for(biography_i in seq_along(biographies_list)) {
  biography <- biographies_list[[biography_i]]

  items_to_fill <- biography |>  keep(is.null) |> names()
  for(item in items_to_fill) {
    biography[[item]] <- generators[[item]](biography)
    biographies_list[[biography_i]] <- biography
    write_yaml(biographies_list, "biographies_list.yaml") 
  }
}
```