### 2023-03-23 code/04-API-interaction-python.qmd

---
title: "API Interaction Python"
format: html
---

## Purpose

Just get the python API working

```{python}
import os
import openai
openai.api_key = os.getenv("OPENAI_API_KEY")

import yaml

with open("config.yml", "r") as ymlfile:
    cfg = yaml.load(ymlfile)

for section in cfg:
    print(section)
print(cfg["default"])

completion = openai.ChatCompletion.create(
  model="gpt-3.5-turbo",
  messages=[
    {"role": "user", "content": "Tell the world about the ChatGPT API in the style of a pirate."}
  ]
)

print(completion.choices[0].message.content)
```