### 2023-03-26 code/04-API-interaction-python.qmd
add pickled file

---
title: "API Interaction Python"
format: html
---

## Purpose

Just get the python API working

### Setup

```{python}
import os
import openai
import pandas as pd
import yaml
import json

with open("config.yml", "r") as f:
    config = yaml.safe_load(f)

openai.api_key = config["default"]["key"]

```

### Translation

```{python}
# biography = "Lundh, Asta Evilla, med. lic., leg. tandläkare, Sthlm, f. i Bunkeflo, Malmöh. l., av lär. Hans L. o. Hanna L. G. 31 m. Waldemar Samuelsson. Barn : Kjell f. 32. Stud:ex. i Malmö 13, med. kand. 19 i Lund, med. lic. 16 i Sthlm, tandlrex. i Sthlm 23. Prakt, läkare o. tandl. i Sthlm sed. 26, tandläkare v. Sthlms folksk. sed. 27."
# 
# translate_prompt = "Translate the following historical biography into English. It includes many Swedish abbreviations:"
# 
# prompt = f"{translate_prompt} \n {biography}"
# 
# completion = openai.ChatCompletion.create(
#   model="gpt-3.5-turbo",
#   messages=[
#     {"role": "user", "content": prompt}
#   ]
# )
# 
# print(completion.choices[0].message.content)
# 
# # remove spaces with strip
# biography_en = completion.choices[0].message.content.strip()
```

```{python}
biography = "Lundh, Asta Evilla, med. lic., leg. tandläkare, Sthlm, f. i Bunkeflo, Malmöh. l., av lär. Hans L. o. Hanna L. G. 31 m. Waldemar Samuelsson. Barn : Kjell f. 32. Stud:ex. i Malmö 13, med. kand. 19 i Lund, med. lic. 16 i Sthlm, tandlrex. i Sthlm 23. Prakt, läkare o. tandl. i Sthlm sed. 26, tandläkare v. Sthlms folksk. sed. 27."

translate_prompt = "Translate the following historical biography into English. It includes many Swedish abbreviations:"
```

```{python}
def translate_text(text):
  print("Going...\n")
  translation = openai.ChatCompletion.create(
    model="gpt-3.5-turbo",
    messages=[
      {"role": "user", "content": f"{translate_prompt}\n\n{text}\n"}
    ]
  ).choices[0].message.content.strip()
  return translation

```

```{python}
biographies = pd.read_excel("data/test_50.xlsx", nrows = 2)
biographies['biography'] = biographies['biography'].replace('\n', ' ', regex=True)
```

```{python}
biographies["biography_translated"] = biographies["biography"].apply(translate_text)

biographies.to_excel("biographies_translated.xlsx", index=False)

```

### Augmentation

schema.org/Person

```{python}
schema_prompt = "Convert this biography into schema.org/Person format.\n* Include information on birthDate, birthPlace, education, occupation, family, parents.\nReturn any other notable information in a field called 'description' where possible.\n* Include startDate and endDate wherever possible.\n* Only provide a  RFC8259 compliant JSON response"

prompt = f"{schema_prompt} \n\n {biography_en}"

completion = openai.ChatCompletion.create(
  model="gpt-3.5-turbo",
  messages=[
    {"role": "system", "content": "You are an expert on Swedish family history."},

    {"role": "user", "content": prompt}
  ]
)

print(completion.choices[0].message.content)

```

```{python}
schema_prompt = "Convert this biography into schema.org/Person format.\n* Return the output in a JSON codeblock\n* Include information on birthDate, birthPlace, education, occupation, family, parents.\nReturn any other notable information in a field called 'description' where possible.\n* Include startDate and endDate wherever possible."

def augment_schema(text):
  print(f"Augmenting {text[:100]}\n")
  augmnetation = openai.ChatCompletion.create(
    model="gpt-3.5-turbo",
    messages=[
      {"role": "user", "content": f"{schema_prompt}\n\n{text}\n"}
    ]
  ).choices[0].message.content
  return augmnetation

```

```{python}
biographies["biography_augmented"] = biographies["biography_translated"].apply(augment_schema)

biographies.to_json("data/biographies.json", orient='records')

print(biographies["biography_augmented"][1])

```

```{python}
biographies['biography_augmented'] = biographies['biography_augmented'].replace('```json', ' ', regex=True)

biographies['biography_augmented'] = biographies['biography_augmented'].replace('```', ' ', regex=True)

biographies['biography_augmented_no_spaces'] = biographies['biography_augmented'].replace('\n', '', regex=True)

biographies['biography_augmented_json'] = biographies['biography_augmented_no_spaces'].apply(lambda x: pd.read_json(x))

```

Occupation

```{python}
occupation_prompt = "Suggest 3 ISCO-08 occupational classifications and 3 O*NET occupational classifications for the indivial descrbied. Return your answer in a JSON codeblock"

prompt = f"{occupation_prompt} \n\n {biography_en}"

completion = openai.ChatCompletion.create(
  model="gpt-3.5-turbo",
  messages=[
    {"role": "user", "content": prompt}
  ]
)

print(completion.choices[0].message.content)

```

```{python}
occupation_prompt = "Suggest 3 ISCO-08 occupational classifications and 3 O*NET occupational classifications for the indivial descrbied. Return your answer in a JSON codeblock"

def augment_occupation(text):
  print(f"Augmenting {text[:100]}\n")
  augmnetation = openai.ChatCompletion.create(
    model="gpt-3.5-turbo",
    messages=[
      {"role": "user", "content": f"{occupation_prompt}\n\n{text}\n"}
    ]
  ).choices[0].message.content
  return augmnetation

```

```{python}
biographies["occupation_augmented"] = biographies["biography_translated"].apply(augment_occupation)

biographies.to_excel("biographies_translated.xlsx", index=False)

```

## New augmenter

```{python}
schema_prompt = "Convert this biography into schema.org/Person format.\n* Include information on birthDate, birthPlace, education, occupation, family, parents where possible.\n * Return any other notable information in a field called 'description' where possible.\n* Include startDate and endDate wherever possible.\n* Only provide a  RFC8259 compliant JSON response"

def augment_biography(text):
  print(f"Augmenting {text[:100]}\n")
  augmnetation = openai.ChatCompletion.create(
    model="gpt-3.5-turbo",
    messages=[
      {"role": "system", "content": "You are an expert on Swedish family history."},

      {"role": "user", "content": f"{schema_prompt}\n\n{text}\n"}
    ]
  ).choices[0].message.content
  return augmnetation

```

```{python}
biography_en = "Lundahl, Karl Johan Theodor J:son, Director, Stockholm, born on November 8, 1877 in Vetlanda to Preacher Samuel Johansson and Amanda J. G. Married to Ruth Ekman for 19 years. After completing his studies at a secondary school, he graduated with distinction from Gothenburg's School of Commerce in 1894. He is a co-owner of John Bernström & Co. in Stockholm since 1908 and has been a director at A-B Ahlsell & Bernström in Stockholm since 1922. He was awarded the Royal Order of Vasa."

test = augment_biography(biography_en)

json.loads(test)

```