### 2023-03-26 code/05-API-interaction-python-2.qmd

---
title: "Python interaction 2"
format: html
---

## Purpose

Translate and Augment biograpies

### Setup

```{python}
import os
import openai
import pandas as pd
import yaml
import json

with open("config.yml", "r") as f:
    config = yaml.safe_load(f)

openai.api_key = config["default"]["key"]
```

### Ingest

```{python}
biographies = pd.read_excel("data/test_50.xlsx", nrows = 5)
biographies['biography'] = biographies['biography'].replace('\n', ' ', regex=True)
```

### Translate

```{python}
translate_prompt = "Translate the following historical biography into English. It includes many Swedish abbreviations:"

def translate_biography(text):
    print(f"Translating {text[:100]}\n")
    translation = None

    try:
        response = openai.ChatCompletion.create(
            model="gpt-3.5-turbo",
            messages=[
                {"role": "system", "content": "You are an expert on Swedish family history."},
                {"role": "user", "content": f"{translate_prompt}\n\n{text}\n"}
            ]
        )
    except openai.OpenAIError as e:
        print(f"Error while creating translation: {e}")
    else:
        if response and response.choices and response.choices[0].message:
            translation = response.choices[0].message.content
        else:
            print("Error: No translation was returned.")

    return translation

```

```{python}
biographies["biography_translated"] = biographies["biography"].apply(translate_biography)

```

### Augment

```{python}
schema_prompt = """Convert this biography into schema.org/Person format.
* Include information on birthDate, birthPlace, education, occupation, family, parents where possible.
* Return any other notable information in a field called 'description' where possible.
* Include startDate and endDate wherever possible.
* Only provide a  RFC8259 compliant JSON response"""

def augment_biography(text):
    print(f"Augmenting {text[:100]}\n")
    try:
        augmentation = openai.ChatCompletion.create(
            model="gpt-3.5-turbo",
            messages=[
                {"role": "system", "content": "You are an expert on Swedish family history."},
                {"role": "user", "content": f"{schema_prompt}\n\n{text}\n"}
            ]
        ).choices[0].message.content

        # Parse the response to a JSON object
        json_obj = json.loads(augmentation)

        return json_obj
    except Exception as e:
        print(f"Error augmenting biography: {e}")
        return None

# Assuming your DataFrame is named `df`
biographies['biography_en_augmented'] = biographies['biography_translated'].apply(augment_biography)

pd.to_pickle(biographies, "data/biographies_augmented.pkl")
```