### 2023-03-28 code/05-API-interaction-python-2.qmd
job data too

---
title: "Python interaction 2"
format: html
---

## Purpose

Translate and Augment biograpies

### Setup

```{python}
import os
import openai
import pandas as pd
import yaml
import json

with open("config.yml", "r") as f:
    config = yaml.safe_load(f)

openai.api_key = config["default"]["key"]
```

### Ingest

```{python}
biographies = pd.read_excel("data/test_50.xlsx")
biographies['biography'] = biographies['biography'].replace('\n', ' ', regex=True)
```

### Translate

```{python}
translate_prompt = "Translate the following historical biography into English. It includes many Swedish abbreviations:"

def translate_biography(text):
    print(f"Translating {text[:100]}\n")
    translation = None

    try:
        response = openai.ChatCompletion.create(
            model="gpt-3.5-turbo",
            messages=[
                {"role": "system", "content": "You are an expert on Swedish family history."},
                {"role": "user", "content": f"{translate_prompt}\n\n{text}\n"}
            ]
        )
    except openai.OpenAIError as e:
        print(f"Error while creating translation: {e}")
    else:
        if response and response.choices and response.choices[0].message:
            translation = response.choices[0].message.content
        else:
            print("Error: No translation was returned.")

    return translation

```

```{python}
biographies["biography_translated"] = biographies["biography"].apply(translate_biography)

pd.to_pickle(biographies, "data/biographies_translated.pkl")

```

### Augment

```{python}
schema_prompt = """Convert this biography into schema.org/Person format.
* Include information on birthDate, birthPlace, education, occupation, family, parents where possible.
* Return any other notable information in a field called 'description' where possible.
* Include startDate and endDate wherever possible.
* Only provide a  RFC8259 compliant JSON response"""

def augment_biography(text):
    print(f"Augmenting {text[:100]}\n")
    try:
        augmentation = openai.ChatCompletion.create(
            model="gpt-3.5-turbo",
            messages=[
                {"role": "system", "content": "You are an expert on Swedish family history."},
                {"role": "user", "content": f"{schema_prompt}\n\n{text}\n"}
            ]
        ).choices[0].message.content

        # Parse the response to a JSON object
        json_obj = json.loads(augmentation)

        return json_obj
    except Exception as e:
        print(f"Error augmenting biography: {e}")
        return None

```

Apply the function

```{python}
# Assuming your DataFrame is named biographies
biographies['biography_en_augmented'] = None  # Create a new column with None values

# Iterate through the DataFrame rows using iterrows()
for index, row in biographies.iterrows():
    # Call the augment_biography function with the biography_translated value
    augmented_biography = augment_biography(row['biography_translated'])

    # Update the biography_en_augmented value in the DataFrame
    biographies.at[index, 'biography_en_augmented'] = augmented_biography

pd.to_pickle(biographies, "data/biographies_augmented.pkl")

```

### Augment education

```{python}
education_prompt = """Read this biography. 
* Classify the highest level of education the individual recieved and the institution they recieved it from
* Return it in JSON where there are two keys: highest_educ_level and educ_institution
* Only provide a  RFC8259 compliant JSON response"""

def augment_education(text):
    print(f"Augmenting education for {text[:100]}\n")
    try:
        augmentation = openai.ChatCompletion.create(
            model="gpt-3.5-turbo",
            messages=[
                {"role": "system", "content": "You are an expert on Swedish family history."},
                {"role": "user", "content": f"{education_prompt}\n\n{text}\n"}
            ]
        ).choices[0].message.content

        # Parse the response to a JSON object
        json_obj = json.loads(augmentation)

        return json_obj
    except Exception as e:
        print(f"Error augmenting education: {e}")
        return None

```

Apply the function

```{python}
# Assuming your DataFrame is named biographies
biographies['education'] = None  # Create a new column with None values

# Iterate through the DataFrame rows using iterrows()
for index, row in biographies.iterrows():
    # Call the augment_education function with the biography_translated value
    augmented_education = augment_education(row['biography_translated'])

    # Update the education value in the DataFrame
    biographies.at[index, 'education'] = augmented_education

pd.to_pickle(biographies, "data/biographies_with_educ_augmented.pkl")

```

### Augment occupation

```{python}
occupation_prompt = """Based on the biography
* Suggest 3 ISCO-08 occupational classifications with a key for code and and a key for description
* Suggest 3 O*NET occupational classifications with a key for code and and a key for description
* Return your answer in JSON format
* Only provide a  RFC8259 compliant JSON response"""

def augment(text, prompt):
    print(f"Augmenting {prompt.split()[0].lower()} for {text[:100]}\n")
    try:
        augmentation = openai.ChatCompletion.create(
            model="gpt-3.5-turbo",
            messages=[
                {"role": "system", "content": "You are an expert on Swedish family history."},
                {"role": "user", "content": f"{prompt}\n\n{text}\n"}
            ]
        ).choices[0].message.content

        # Parse the response to a JSON object
        json_obj = json.loads(augmentation)

        return json_obj
    except Exception as e:
        print(f"Error augmenting {prompt.split()[0].lower()}: {e}")
        return None

```

Apply the function

```{python}
# Assuming your DataFrame is named biographies
biographies['occupation'] = None  # Create a new column with None values

# Iterate through the DataFrame rows using iterrows()
for index, row in biographies.iterrows():
    # Call the augment_education function with the biography_translated value
    augmented_occupation = augment(row['biography_translated'], occupation_prompt)

    # Update the occupation value in the DataFrame
    biographies.at[index, 'occupation'] = augmented_occupation
    print(f"Occupation for row {index}: {augmented_occupation}")

pd.to_pickle(biographies, "data/biographies_with_educ_and_occupation_augmented.pkl")

```

## Export data

```{python}
biographies.to_excel("data/50_test_complete.xlsx")

```