### 2023-03-28 code/06-API-interaction-python-3.qmd

---
title: "06-API interaction"
format: html
---

```{python}
def augment(text, prompt):
    print(f"Augmenting {prompt.split()[0].lower()} for {text[:100]}\n")
    try:
        augmentation = openai.ChatCompletion.create(
            model="gpt-3.5-turbo",
            messages=[
                {"role": "system", "content": "You are an expert on Swedish family history."},
                {"role": "user", "content": f"{prompt}\n\n{text}\n"}
            ]
        ).choices[0].message.content

        # Parse the response to a JSON object
        json_obj = json.loads(augmentation)

        return json_obj
    except Exception as e:
        print(f"Error augmenting {prompt.split()[0].lower()}: {e}")
        return None

# Then, you can call the function with the appropriate prompt
biography_augmented = augment(text, schema_prompt)
education_augmented = augment(text, education_prompt)

```