### 2023-04-11 code/07-Line-splitter.qmd
add scraper

---
title: "Splitting the text"
format: html
---

## Purpose

I need to split the text up that I scrape.

### Prompt

Hello GPT-4. I am trying to solve a data science problem, can you give me a hand? I've heard that you're an expert strategic thinker and coder.

I have scraped a series of biographies of individuals from a website. The problem I have is that the biographies are not broken up by any identifiable token. The data I have are lines of text and I want to group all of the lines of text that belong to the same biography together. There are multiple biographies per web page, between 4 and 8. The biographies are in alphabetical order. For instance, I have a biography for "Abrahamson, Ivar Magnus" followed by one for "Abramson, Ernst Leopold". I was thinking of trying to train an LSTM to identify breaks in my series of text just before a new name. Do you think this is the best way of doing this?

## Response

It suggests using regex. Probably good enough.

```{python}
import re

def split_biographies(lines):
    biographies = []
    current_bio = []

    # Regex pattern to match a name in the format "Lastname, Firstname(s) (additional information)"
    name_pattern = re.compile(r'^[A-Za-z]+(?:-[A-Za-z]+)?,\s[A-Za-z]+(?:\s[A-Za-z]+)*(?:\s[A-Za-z.]+)*,')

    for line in lines:
        if name_pattern.match(line):
            if current_bio:  # Save the previous biography if it's not empty
                biographies.append(current_bio)
                current_bio = []
        current_bio.append(line)

    # Save the last biography
    if current_bio:
        biographies.append(current_bio)

    return biographies

# Example usage
lines = [
    "Abrahamson, Ivar Magnus, prakt,",
    "...",
    "Acking, Carl-Axel, arkitekt, Sthlm,",
    "...",
]

biographies = split_biographies(lines)

```

```{python}

```