### 2023-05-14 presentations/presentation-CBS-14-05-2023.qmd
add pres

---
title: "Wired for Success: Sweden's Electrical Innovators in the Age of Industrial Transformation"
title-slide-attributes:
  data-background-image: "assets/preview.png"
  data-background-size: "30%"
  data-background-opacity: "0.6"
format: 
  revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://j-jayes.github.io/Data-Visualization-in-R-Lund/)"
image: "preview_img/03-coordinate-systems-and-axes.jfif"
favicon: html/favicon.ico
---

## Motivation

1. Electrification is a cannonical example of a **General Purpose Technology (GPT)**

2. When this new technology was adopted in Sweden:

    - It made individuals richer
    - It lowered inequality
    - It allowed structural change on the labour market

2. What we want to know next:

    - Who were the innovators or adopters?
    - Did they move to opportunity?
    - How much training did they have in the new technology?

## What I want you to take away from today as researchers

:::: {.columns}

::: {.column width="40%"}
1. NLP tasks that used to be very challenging / time consuming are no longer!

2. We can use the OpenAI API to structure data from free text at very low cost.

3. It is a General Purpose Technology that can both process data and also inform our workflow as researchers.

:::

::: {.column width="10%"}
:::

::: {.column width="50%"}
![](assets/robot.png){fig-align="center" width=60%}
:::

::::

## Why do we care about the second industrial revolution?

![](assets/factory.png){fig-align="center"}

##

![](assets/2IR_1.JPG){fig-align="center"}

##

![](assets/2IR_2.JPG){fig-align="center"}

## New technology with rapid adoption in Sweden

![](assets/schon-energy-share.jpeg){fig-align="center"}

## What happened to early electrifiers in Sweden?

![](assets/grid_rollout.jpg){fig-align="center"}

## No large unemployment effects in areas who electrified first

![](assets/employment_share.jpeg){fig-align="center"}

## Incomes were markedly higher in areas that electrified first

![](assets/coef_plot.jpeg){fig-align="center"}

## Income was shared more equally in areas that electrified first

![](assets/dot_plot.jpeg){fig-align="center"}

## Now we want to ask:

- Who were the innovators or adopters?
- Did they move to opportunity?
- How much training did they have in the new technology?

## Who is who?

::: {.columns}

::: {.column width="70%"}

Vem är Vem? is a **Swedish biographical encyclopedia** that was published in two editions of five volumes each in 1945–1950 and 1962–1968 by Bokförlaget Vem är Vem.

The intention was, according to the publishers, to draw attention **to people who were at the height of their activities**, even if they were younger in influential or otherwise noted positions in different areas.

We get biographies and career trajectories of 69448 individuals!

8 of the 10 volumes are digitized by librarians in Uppsala – thank you <3

:::

::: {.column width="10%"}

:::

::: {.column width="20%"}

![](assets/vem_cover.png){fig-align="center" width=100%}

:::

## Vem är Vem?

```{r}
library(tidyverse)
theme_set(theme_light())

# Create a tibble from the JSON data
books <- tibble(
  book_name = c("Skåne", "Stockholm", "Svealand", "The Götaland region except Skåne", "Skåne", "Norrland", "Stockholm", "The Götaland region except Skåne"),
  year = c(1948, 1962, 1964, 1965, 1966, 1968, 1945, 1948),
  pages = c(624, 1498, 945, 1203, 952, 1340, 1044, 1075),
  volume = c(1, 2, 2, 2, 2, 2, 1, 1)
)

books <- books %>%
    mutate( name = str_c(book_name, " - ", year),
            name = fct_reorder(name, pages),
            records = pages * 8) %>%
            mutate(volume = if_else(volume == 1, "1945–1950", "1962–1968")) %>%
    select(name, records, volume)

# books %>% summarise(records = sum(records))

# Create the ggplot visualization
ggplot(books, aes(x = records, y = name, fill = factor(volume))) +
  geom_col(alpha = .8) +
  labs(x = "Number of biographies", y = "Book volume",
  title = "Number of biographies in Vem är Vem?",
  fill = "Edition") +
  scale_x_continuous(labels = scales::number_format()) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800"))

```

## Vem är Vem? Example

![](assets/Karl_Lund.png)

## Vem är Karl Lund?

Lund, Karl Gustaf, chief engineer, Varberg, born on July 22, 1893, in Hille, Gävle County, Sweden, son of factory worker Ferdinand L. and Maria Andersson. Married in 1936 to Sigrid Johansson. Children: Ingvar (born 1938), Lennart (born 1942). — Graduated from Bergsskolan in Filipstad in 1917, specialized studies at the Royal Institute of Technology (KTH) from 1920 to 1922, studied at the Institute of Metallurgy and Stockholm University in 1921-1922. Chemist at Strömsnäs Järnverks A-B in Degerfors from 1918 to 1920, metallurgist and chemist at Westinghouse Electric & Manufacturing Co. in East Pittsburgh, PA, USA, from 1923 to 1926 and 1928 to 1929, chief metallurgist at Laclede Steel Co. in Alton, Illinois, USA, in 1927, furnace and steelworks engineer at A-B Iggesunds Bruk from 1929 to 1931, site manager at Gunnebo Bruks Nya A-B, Varbergsverket, since 1931. Member of the municipal executive committee, deputy chairman of the economic department, deputy member of the board of the power plant, chairman of Varbergs Sparbank, employer representative in the district council of the county labor board, member of the board of Varbergs Luftskyddsförening (Varberg Air Protection Association), secretary of Varbergs Högerförening (Varberg Conservative Association), chairman of the railway sick fund, and Plant Society for Small Bird Friends.

## Vem är Karl Lund?

Traveled to Germany in 1921, 1923, 1930, and 1936, Denmark, Czechoslovakia in 1921, 1922, and 1923, Austria in 1921, and the USA from 1923 to 1929. Writings: "Some fundamental factors for obtaining sharp thermal curves" (Transactions of the American Society for Steel Treating, co-authored with C. Benedicks and W. H. Dearden, 1925), "Contemporary production of saw blades, circular saws, and machine knives" (Timber Industry, 1931). Hobbies: hunting and fishing.

## Vem är Karl Lund?

[Lund, Karl Gustaf]{style="background-color: #b0c4de"}, [chief engineer]{style="background-color: #d2b48c"}, [born on July 22, 1893]{style="background-color: #ff8c69"}, [in Hille, Gävle County]{style="background-color: #90ee90"}, Sweden, son of [factory worker Ferdinand L. and Maria Andersson]{style="background-color: #9370db"}. Married in 1936 to [Sigrid Johansson]{style="background-color: #ffadc1"}. Children: [Ingvar (born 1938)]{style="background-color: #ffdead"}, [Lennart (born 1942)]{style="background-color: #ffdead"}. — Graduated from [Bergsskolan]{style="background-color: #afeeee"} in Filipstad in 1917, specialized studies at the [Royal Institute of Technology (KTH)]{style="background-color: #afeeee"} from 1920 to 1922, studied at the [Institute of Metallurgy]{style="background-color: #afeeee"} and [Stockholm University]{style="background-color: #afeeee"} in 1921-1922. Chemist at [Strömsnäs Järnverks A-B]{style="background-color: #98fb98"} in Degerfors from 1918 to 1920, metallurgist and chemist at [Westinghouse Electric & Manufacturing Co.]{style="background-color: #98fb98"} in East Pittsburgh, PA, USA, from 1923 to 1926 and 1928 to 1929, chief metallurgist at [Laclede Steel Co.]{style="background-color: #98fb98"} in Alton, Illinois, USA, in 1927, furnace and steelworks engineer at [A-B Iggesunds Bruk]{style="background-color: #98fb98"} from 1929 to 1931, site manager at [Gunnebo Bruks Nya A-B, Varbergsverket]{style="background-color: #98fb98"}, since 1931. Member of the municipal executive committee, deputy chairman of the economic department, deputy member of the board of the power plant, chairman of [Varbergs Sparbank]{style="background-color: #ffdab9"}, employer representative in the district council of the county labor board, member of the board of [Varbergs Luftskyddsförening (Varberg Air Protection Association)]{style="background-color: #ffdab9"}, secretary of [Varbergs Högerförening (Varberg Conservative Association)]{style="background-color: #ffdab9"}, chairman of the railway sick fund, and [Plant Society for Small Bird Friends]{style="background-color: #ffdab9"}.

## Vem är Karl Lund?

Traveled to [Germany]{style="background-color: #e6e6fa"} in 1921, 1923, 1930, and 1936, [Denmark]{style="background-color: #e6e6fa"}, [Czechoslovakia]{style="background-color: #e6e6fa"} in 1921, 1922, and 1923, [Austria]{style="background-color: #e6e6fa"} in 1921, and the [USA]{style="background-color: #e6e6fa"} from 1923 to 1929. Writings: "Some fundamental factors for obtaining sharp thermal curves" (Transactions of the American Society for Steel Treating, co-authored with C. Benedicks and W. H. Dearden, 1925), "Contemporary production of saw blades, circular saws, and machine knives" (Timber Industry, 1931). Hobbies: [hunting]{style="background-color: #ffd700"} and [fishing]{style="background-color: #ffd700"}.

## Birthplaces mapped

```{r}
knitr::include_url("assets/sweden_map_checkboxes.html", height = "800")
```

## How were NLP tasks hard before?

In your case when you use patent data, a lot of the research on historical questions is done just with the title of the patent and relationships between patent holders. It was difficult to get more information from the text of the patent.

## How are they easy now?

Now we can use the OCR / NLP OpenAI API to get a lot of information from the text of the patent.