### 2024-01-10 template.qmd

---
title: |
  On the Surgical Removal of \
  Cardassian Cranial Implants
subtitle: A case study of a Cardassian patient
short-title: Cranial Implants
published: "Under review at *Starfleet Medical Journal*"
code-repo: "Access the code, data, and analysis at <https://git.starfleet.ufp/bashir/cranial-implants>"
correspondence-prefix: "Correspondence concerning this article should be addressed to"
author:
  - name: Julian Bashir
    email: jbashir@starfleet.ufp
    orcid: 0000-0002-3948-3914
    title: Chief Medical Officer
    affiliations: 
      - id: ds9
        name: Starbase Deep Space Nine
        department: Sick Bay
        address: 1234 Main Street
        city: Anytown
        region: NY
        country: USA
        postal-code: 90210
    attributes:
      corresponding: true
  - name: Elim Garak
    email: tailor.spy@obsidianorder.card.gov
    title: Shopkeeper and Tailor
    affiliations:
      - id: terok
        name: Terok Nor
        department: Promenade
      - ref: ds9
abstract: |
  Space: the final frontier. These are the voyages of the starship *Enterprise*. Its continuing mission: to explore strange new worlds. To seek out new life and new civilizations. To boldly go where no one has gone before!
thanks: |
  Placeholder text generated at <https://vlad-saling.github.io/star-trek-ipsum/>.
additional-info: |
  We have no known conflict of interest to disclose.
keywords:
  - surgery
  - espionage
  - brains
date: August 24, 2022
epigraph:
  - text: War is good for business. 
    source: Ferengi Rule of Aquisition 34
  - text: Peace is good for business. 
    source: Ferengi Rule of Aquisition 35
bibliography: references.bib
format:
  hikmah-pdf: default
  hikmah-manuscript-pdf: default
---

We're acquainted with the wormhole phenomenon, but this... Is a remarkable piece of bio-electronic engineering by which I see much of the EM spectrum ranging from heat and infrared through radio waves, et cetera, and forgive me if I've said and listened to this a thousand times [@Lovelace:1842]. This planet's interior heat provides an abundance of geothermal energy. We need to neutralize the homing signal.

# Main section

It indicates a synchronic distortion in the areas emanating triolic waves. The cerebellum, the cerebral cortex, the brain stem,  the entire nervous system has been depleted of electrochemical energy [@Turing:1936].

```{r}
n_lights <- 2 + 2
n_lights
```

Any device like that would produce high levels of triolic waves. These walls have undergone some kind of selective molecular polarization. I haven't determined if our phaser energy can generate a stable field. We could alter the photons with phase discriminators.

$$
y = \text{Something} + \beta_1 x_1
$$

## Subsection

Communication is not possible [@Keynes:1937]. The shuttle has no power (see @fig-n-lights).

Using the gravitational pull of a star to slingshot back in time? We are going to Starbase Montgomery for Engineering consultations prompted by minor read-out anomalies. Probes have recorded unusual levels of geological activity in all five planetary systems.

Assemble a team. Look at records of the Drema quadrant. Would these scans detect artificial transmissions as well as natural signals?

```{r plot-n-lights}
#| echo: false
#| label: fig-n-lights
#| fig-cap: There are four lights
#| fig-width: 5
#| fig-height: 1
#| fig-align: "center"
par(mar = c(0, 0, 0, 0))
plot(x = seq(1:n_lights), y = rep(1, 4), col = "darkgoldenrod1", 
     pch = 8, cex = 4, axes = TRUE, ann = FALSE, xlab = FALSE, ylab = FALSE)
```

## Other heading

Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo

### Third

inventore veritatis^[Thing] et quasi architecto beatae vitae dicta sunt explicabo. Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt. Neque porro quisquam est,^[Another thing]

qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem.

#### A 4th level heading

qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem.