### 2024-01-11 table-process.qmd

---
title: "table"
format: html
---

```{r}
library(tidyverse)
library(gt)

data <- tribble(
  ~Step, ~Process,                               ~Icons,   ~Description,
  1,     "Scrape book data from website",         "🌐➡️📚",  "",
  2,     "Split records on each page of a book", "📚➡️📄", "",
  3,     "Structure records with LLM",                 "📄🔀📄",   "",
  4,     "Augment data with coordinates",                          "🔗🔍🌍🎓📚", "",
  5,     "Store data for analysis",                            "🔗📚📂",   ""
)

data %>%
select(-Description) %>% 
gt() %>% 
cols_label(Icons = "") %>% 
tab_header(title = md("**Data Collection Strategy**"),
subtitle = "For biographical dictionaries and industry catalogue") %>% 
gtsave("assets/data-collection-process.png")

```