### 2024-01-22 index.qmd
add presi

---
title: "Praise the people or praise the place? Upper tail human capital in electrifying Sweden"
subtitle: "Midway Seminar - 24-01-2024"
author: 
  - "Jonathan Jayes  - Lund University"
title-slide-attributes:
  data-background-image: "assets/preview_wide.png"
  data-background-position: "bottom 10"
  data-background-size: "90%"
  data-background-opacity: "0.4"
bibliography: bibliography.bib
format: 
  revealjs:
    code-overflow: wrap
    width: 1600
    height: 900
    fig-align: center
    default-image-extension: svg
    footer: "[{{< fa house >}}](https://interludeone.com/content/research.html)"
    slide-number: c
    code-block-bg: true
    code-block-border-left: "#31BAE9"
    include-in-header: 
        - text: <link rel = "shortcut icon" href = "assets/favicon.ico" />

---

# Research question{#sec-researchquestion background-color="#1C9E77" visibility="uncounted"}

<br>

How can we use new kinds of source material and new methods to find out where the skilled workers who built the Swedish electricity sector came from, and where they were trained?

Answering involves:

- 2 data structuring tasks
- 2 classification tasks

## Motivation

1. Electrification is a cannonical example of a **General Purpose Technology (GPT)**

2. When this new technology was adopted in Sweden:

    - It made individuals richer 💰🔋
    - It lowered inequality 📉🤝
    - It allowed structural change on the labor market 🔧💼
    - Middle-skill workers were drawn from the local area to the new jobs 🏙🏃‍♀️

2. What we want to know next:

    - Who were the innovators or adopters? 🤔🚀
    - Did they move to opportunity? 🏃‍♀️💼
    - How much training did they have in the new technology? 🎓⚡️
    - Were did they get this training? 🏫🔌

## Related literature

**3 main strands:**

::: {.columns}
::: {.column}
*Recent papers making use of individual level biographic data*

- @ford2023not
- @Nekoei2020Herstory

:::
::: {.column}
*The importance of human capital in economic development*

- @Goldin1994
- @Moretti2012
- @Mokyr2017

:::
::: {.column}
*Classification and prediction as part of EH*

- @DahlVedel2024
- @merouani2023innovation
- @Bandiera2020ceo

:::
:::

# Who were the `innovators` or adopters?{#sec-who background-color="#1C9E77" visibility="uncounted"}

## Who is who?

:::: {.columns}

::: {.column width="70%"}

Vem är Vem? is a **Swedish biographical encyclopedia** that was published in two editions of five volumes each in 1945–1950 and 1962–1968 by Bokförlaget Vem är Vem.

The intention was, according to the publishers, to draw attention **to people who were at the height of their activities**, even if they were younger, in influential or otherwise noted positions in different areas.

- Biographies and career trajectories of ~ 75,000 individuals!

- 9 of the 11 volumes are digitized by librarians in Uppsala – thank you <3

:::

::: {.column width="5%"}

:::

::: {.column width="25%"}

![](assets/vem_page.png){fig-align="center" width=100%}

:::

::::

## Vem är Vem?

```{r}
library(tidyverse)
theme_set(theme_light())

# Create a tibble from the JSON data
books <- tibble(
    book_name = c("Skåne", "Stockholm", "Svealand", "The Götaland region except Skåne", "Skåne", "Norrland", "Stockholm", "The Götaland region except Skåne", "Who is who in indsutry and business?"),
    year = c(1948, 1962, 1964, 1965, 1966, 1968, 1945, 1948, 1945),
    pages = c(624, 1498, 945, 1203, 952, 1340, 1044, 1075, 625),
    volume = c(1, 2, 2, 2, 2, 2, 1, 1, 3)
)

books <- books %>%
    mutate(
        name = str_c(book_name, " - ", year),
        name = fct_reorder(name, pages),
        records = pages * 8
    ) %>%
    mutate(volume = case_when(
        volume == 1 ~ "1945–1950",
        volume == 2 ~ "1962–1968",
        TRUE ~ "Who is who in indsutry and business?"
    )) %>%
    select(name, records, volume)

# books %>% summarise(records = sum(records))

# Create the ggplot visualization
ggplot(books, aes(x = records, y = name, fill = factor(volume))) +
    geom_col(alpha = .8) +
    labs(
        x = "Number of biographies", y = "Book volume",
        title = "Number of biographies in Vem är Vem?",
        fill = "Edition"
    ) +
    scale_x_continuous(labels = scales::number_format()) +
    scale_fill_manual(values = c("#2B8CBE", "#be5d2b", "midnightblue")) +
    theme(plot.title.position = "plot")

# ggsave("assets/who-is-who-volumes.png", width = 20, height = 20, units = "cm")

```

## Vem är Karl Lund?

[Lund, Karl Gustaf]{style="background-color: #b0c4de"}, [chief engineer]{style="background-color: #d2b48c"}, [born on July 22, 1893]{style="background-color: #ff8c69"}, [in Hille, Gävleborgs län]{style="background-color: #90ee90"}, Sweden, son of [clerk Ferdinand L. and Maria Andersson]{style="background-color: #9370db"}. Married in 1936 to [Sigrid Johansson]{style="background-color: #ffadc1"}. Children: [Ingvar (born 1938)]{style="background-color: #ffdead"}, [Lennart (born 1942)]{style="background-color: #ffdead"}. — Graduated from [Bergsskolan]{style="background-color: #afeeee"} in Filipstad in 1917, specialized studies at the [Royal Institute of Technology (KTH)]{style="background-color: #afeeee"} from 1920 to 1922, studied at the [Institute of Metallurgy]{style="background-color: #afeeee"} and [Stockholm University]{style="background-color: #afeeee"} in 1921-1922. Chemist at [Strömsnäs Järnverks A-B]{style="background-color: #98fb98"} in Degerfors from 1918 to 1920, metallurgist and chemist at [Westinghouse Electric & Manufacturing Co.]{style="background-color: #98fb98"} in East Pittsburgh, PA, USA, from 1923 to 1926 and 1928 to 1929, chief metallurgist at [Laclede Steel Co.]{style="background-color: #98fb98"} in Alton, Illinois, USA, in 1927, furnace and steelworks engineer at [A-B Iggesunds Bruk]{style="background-color: #98fb98"} from 1929 to 1931, site manager at [Gunnebo Bruks Nya A-B, Varbergsverket]{style="background-color: #98fb98"}, since 1931. Member of the municipal executive committee, deputy chairman of the economic department, deputy member of the board of the power plant, chairman of [Varbergs Sparbank]{style="background-color: #ffdab9"}, employer representative in the district council of the county labor board, member of the board of [Varbergs Luftskyddsförening (Varberg Air Protection Association)]{style="background-color: #ffdab9"}, secretary of [Varbergs Högerförening (Varberg Conservative Association)]{style="background-color: #ffdab9"}, chairman of the railway sick fund, and [Plant Society for Small Bird Friends]{style="background-color: #ffdab9"}.

# Where were the `firms` in which these individuals worked?{#sec-where background-color="#1C9E77" visibility="uncounted"}

## Swedish industrial firm catalogue of 1947

::: {.columns}
::: {.column}
![](assets/svensk-industrikalender_1.png){fig-align="center" width=100%}
:::
::: {.column}
The Svensk Industrikalender, published by Sveriges Industriförbund (Sweden’s Industrial Association), is a comprehensive directory of Swedish industrial firms. This calendar was issued annually from 1918 to 2000 and contains information related to Swedish industry. The 1947 edition available on the Project Runeberg website was digitized in April 2012, sourced from the Centrum för Näringslivshis- toria.
:::
:::

## Firms across the country

```{=html}
<div style="display: flex; justify-content: center; align-items: center;">
    <iframe src="https://raw.githack.com/j-jayes/svensk-industrikalender/main/data/maps/all_companies_map.html" width="1400" height="700"></iframe>
</div>
```

# How can we `structure` the data?{#sec-structure background-color="#1C9E77" visibility="uncounted"}

## Data pipeline

<br>

| Step | Process                                   |        | Description                                                           |
|------|-------------------------------------------|-------------|
| 1    | Scrape book data from website             | 🌐➡️📚       | 
| 2    | Split records on each page of a book  | 📚➡️📄       | 
| **3**    | **Structure biographies using schema**    | 📄🔄📄    | 
| **4**    | **Classify record into group**                              | 🔗🔍🌍🎓📚   |
| 5    | Store data for analysis                                | 🔗📚📂       |

## NLP challenges

:::: {.columns}

::: {.column width="50%"}

<br>

- Many abbreviations and contractions
    - DOB: "f\.\\s*(\\d{2})\\/(\\d{2})/(\\d{2})"
    - Gävleborg County: "Gävleb. l."

- Similar structure for each entry but not exactly the same information in the same order

:::

::: {.column width="50%"}

<br>

![](assets/Karl_Lund_crop.png){fig-align="center"}

:::

::::

## Structure using OpenAI API 🤖

### Using OpenAI's GPT-3.5 chat API:

We provide the system a schema to structure the record in, e.g. keys.

We provide the system the Swedish text

We specify the JSON response format we want

<br>

::: {.columns}
::: {.column}

```{python, eval=FALSE, echo=TRUE}
schema = {
    "type": "object",
    "required": [
        "full_name",
        "location",
        "occupation",
        "birth_details",
        "education",
        "career",
        "family",
    ],
    "properties": {
        "full_name": {"type": "string"},
        "location": {"type": "string"},
        "occupation": {"type": "string"},
        "birth_details": {
            "type": "object",
            "properties": {
                "date": {"type": "string"},
                "place": {"type": "string"},
                "parents": {"type": "string"},
            },
            "required": ["date", "place"],
        },}}"
```

:::
::: {.column}

```{python, eval=FALSE, echo=TRUE}

def structure_biography_info(page_text):
    try:
        # Create a prompt for the system
        structure_prompt = f"Task: read the schema and return RFC compliant JSON information about the Swedish individuals from the 1950 biographical dictionary 'Vem är Vem' that is provided below. Use a numeric index for each biography in your JSON output and return information about all of them, including all career information available. Keep the biographic descriptions in Swedish and remove any abbreviations based on your knowledge, e.g. 'fil. kand.' is 'filosofie kandidat', and 'Skarab. l.' is 'Skaraborgs Län'. Put years in full based on context. Put dates in dd/mm/yyyy format where possible. If there is no information for a key, leave it out. If there is no information for a required key, put NULL as the value.\nHere is the schema: {schema}.\nHere is the text: {page_text}. Go!"
        structure_response = client.chat.completions.create(
            model="gpt-3.5-turbo-1106",
            response_format={"type": "json_object"},
            messages=[
                {
                    "role": "system",
                    "content": "You are an expert on Swedish biographies.",
                },
                {"role": "user", "content": f"{structure_prompt}"},
            ],
        )

        structured_biography_info = json.loads(
            structure_response.choices[0].message.content
        )

        return structured_biography_info

    except Exception as e:
        print(f"Error in structure_biography_info: {e}")
        return None
```

:::
:::

## Add coordinates to firms and individuals

Birthplaces of individuals who studied at Royal Tehcnical Institute (KTH) in *Who is Who in Industry and Business*

<div style="text-align: center">

```{r}
knitr::include_url("assets/kth_birthplace_map.html", height = "800")
```

</div>

# How can we `classify occupations` according to a schema?{#sec-classifying-occupations background-color="#1C9E77" visibility="uncounted"}

```{mermaid}
graph LR
    A[Queried Job Title: Civilingenjör, teknisk fysik Code: 2142] --> B[Text Embeddings]
    B --> C[Classifier Algorithm]
    C --> D[Similarity Ranking]
    D --> E1[1. 2142 - Civilingenjörsyrken inom bygg och anläggning]
    D --> E2[2. 8212 - Montörer, elektrisk och elektronisk utrustning]
    D --> E3[3. 7215 - Stålkonstruktionsmontörer och grovplåtslagare]
    D --> E4[4. 7319 - Musikinstrumentmakare och övriga konsthantverkare]
    D --> E5[5. 1212 - Ekonomi- och finanschefer nivå 2]

    style E1 fill:#ffff
    style A fill:#ffff

```

## Two-part classification process

Makes use of occupational title, e.g. "Civilingenjör, teknisk fysik" and workplace information, e.g. "Strömsnäs Järnverks A-B".

```{python, eval=FALSE, echo=TRUE}
#| code-line-numbers: "10-14"
{
  "career": [
    {
      "occupational_title": "Chemist",
      "workplace": "Strömsnäs Järnverks A-B",
      "industry": "Metallurgy",
      "start_date": "1918"
    },
    {
      "occupational_title": "Metallurgist and Chemist",
      "workplace": "Westinghouse Electric & Manufacturing Co.",
      "industry": "Manufacturing",
      "start_date": "1923"
    },
    {
      "occupational_title": "Chief Metallurgist",
      "workplace": "Laclede Steel Co.",
      "industry": "Steel Production",
      "start_date": "1927"
    },
    {
      "occupational_title": "Furnace and Steelworks Engineer",
      "workplace": "A-B Iggesunds Bruk",
      "industry": "Metallurgy",
      "start_date": "1929"
    },
    {
      "occupational_title": "Site Manager",
      "workplace": "Gunnebo Bruks Nya A-B, Varbergsverket",
      "industry": "Metallurgy",
      "start_date": "1931"
    }
  ]
}

```

## Why was classification hard?

### Fuzzy string matching doesn't capture semantic similarity between words
- "chauffeur" is close to "car driver" semantically, but not close in text.

### Many possible groups makes classification challenging for traditional ML models
- HISCO schema has 1,600 unit groups.

### Building one labelled classification scheme manually does not generalize easily
- If you code all of your occupations to HISCO and then want to move to SSYK, you have to redo all of the codings or try and find a crosswalk.

:::: {.columns}

::: {.column width="40%"}
I spent months doing this last year to classify occupations to HISCLASS groups, hand labelling data and optimizing a support vector machine model.
:::

::: {.column width="60%"}

<div style="text-align: center">

**It did not work**

```{r}
knitr::include_graphics("assets/svm.png")
```
</div>
:::

::::

## Why is classification easier now?

### We make use of pre-trained large language models
- Benefit from semantic similarity
- Text embeddings mean we can have an arbitrary number of classes
- Switch out classification target schema easily!

<div style="text-align: center">

```{r}
knitr::include_graphics("assets/sustainable-planning.png")
```

</div>

## What is an embedding?

![](assets/word2vec-king-queen-vectors.png){fig-align="center" width=100%}

## What is an embedding?{.unnumbered}

![](assets/word2vec-king-queen-composition.png){fig-align="center" width=100%}

## Use embeddings to cluster skills

```{r}
#| label: fig-hisco-cluster
#| fig-cap: "Scatter plot showing the relative similarity of occupational titles in the HISCO schema"

knitr::include_graphics("assets/hisco_cluster.png")
```

## Use clusters to classify individuals into the groups

```{r}
#| label: fig-hisco-counts
#| fig-cap: "Most common HISCO occupations in Vem är Vem"
#| fig-width: 5
#| out-width: 5in

knitr::include_graphics("assets/hisco_counts.png")
```

## Use clusters to classify new skills into the groups

```{mermaid}
graph TB
    A[Collect HISCO Codes<br>and Descriptions] -->|Use OpenAI API| B[Convert HISCO Titles<br>and Descriptions to Vectors]
    C[Receive Occupational<br>Strings] -->|Use OpenAI API| D[Convert Occupational<br>Strings to Vectors]
    B --> E[Compare Vectors in<br>Vector Space]
    D --> E
    E --> F[Assign HISCO Code to<br>Occupational String<br>using Cosine Distance]
    style A fill:#2B8CBE
    style B fill:#be5d2b
    style C fill:#2B8CBE
    style D fill:#be5d2b
    style E fill:#6f6fbf
    style F fill:#F4FA58

```

Here the HISCO descriptions are in English, as are the translated titles from my biographies in Swedish.

## Cluster firms by location

```{r}
#| label: fig-map-common-types
#| fig-cap: "Map of the geographic clusters of businesses by most common business type"
#| fig-width: 5
#| out-width: 5in

knitr::include_graphics("assets/map_most_common_businesses.png")
```

## You can do it too!

![](assets/code.png){fig-align="center" width=100%}

[GitHub repo](https://github.com/j-jayes/who-is-who-in-industry)

## References

::: {#refs}
:::

# Appendix {#sec-appendix background-color="#1C9E77" visibility="uncounted"}

## How to get a text embedding?{visibility="uncounted"}

```{python}
#| eval: false
#| echo: true

# Define a function to get the text embedding
def get_embedding(text, model="text-embedding-ada-002"):
    text = text.replace("\n", " ")
    return openai.Embedding.create(input=[text], model=model)['data'][0]['embedding']

# Apply the get_embedding function to your 'hisco_text' column
df['ada_embedding'] = df['hisco_text'].apply(lambda x: get_embedding(x, model='text-embedding-ada-002'))

```

## Assign HISCOs: example{visibility="uncounted"}

### 026 Metallurgists (HISCO description)

Workers in this unit group advise on metallurgical problems and develop and control processes for the extraction of metals from their ores, study the properties and engineering characteristics of metals and develop new alloys, and develop and supervise metal manufacturing processes for making finished metal products.

### Karl Lund in 1923 (Input)

Metallurgist and Chemist at Westinghouse Electric & Manufacturing Co.

## Assign HISCOs: codes projected from 1536 dimensions to 2 with UMAP{visibility="uncounted"}

```{r}
knitr::include_url("assets/interactive_umap_plot.html", height = "800")
```

## Assign HISCOs: Watchmaker Apprentice{visibility="uncounted"}

<br>

```{r}
library(gt)

# Create a tibble with the provided data
watchmaker_data <- tibble(
    Rank = c(1, 2, 3, 4, 5),
    HISCO_CODE = c(842, 830, 810, 811, 839),
    Occupational_Description = c(
        "Watch, Clock and Precision Instrument Makers",
        "Blacksmiths, Toolmakers and Machine-Tool Operators",
        "Woodworkers",
        "Cabinetmakers",
        "Blacksmiths, Toolmakers and Machine-Tool Operators"
    )
)

# Create a gt table from the tibble
watchmaker_table <- watchmaker_data %>%
    gt() %>%
    tab_header(
        title = md("**Closest HISCO codes** to Watchmaker Apprentice")
    ) %>%
    cols_label(
        Rank = "Rank",
        HISCO_CODE = "HISCO CODE",
        Occupational_Description = "Occupational Description"
    )

watchmaker_table %>%
    tab_options(table.font.size = 40)
```

## Assign HISCOs: Shipyard Assistant Engineer{visibility="uncounted"}

<br>

```{r}

# Create a tibble with the provided data
occupation_data <- tibble(
    Rank = c(1, 2, 3, 4, 5),
    HISCO_CODE = c(43, 24, 982, 42, 959),
    Occupational_Description = c(
        "Ships' Engineers",
        "Mechanical Engineer",
        "Ships' Engine-Room Workers",
        "Ships' Deck Officers and Pilots",
        "Construction Workers"
    )
)

# Create a gt table from the tibble
occupation_table <- occupation_data %>%
    gt() %>%
    tab_header(
        title = md("**Closest HISCO Codes** to 'Shipyard Assistant Engineer'")
    ) %>%
    cols_label(
        Rank = "Rank",
        HISCO_CODE = "HISCO CODE",
        Occupational_Description = "Occupational Description"
    )

occupation_table %>%
    tab_options(table.font.size = 40)

```

## Keep titles in Swedish?{visibility="uncounted"}

Instead of using the HISCO schema in English, I am using the SSYK classification from 2012.

Instead of using the multilingual OpenAI ADA model, I now use [The National Library of Sweden Swedish BERT model](https://huggingface.co/KB/bert-base-swedish-cased). It is trained on 200M sentences, 3000M tokens from books, news, government publications, swedish wikipedia and internet forums.

The input is from the Job Tech Data so we have a ground truth that they have labelled.

```{r}
knitr::include_url("https://huggingface.co/KB/bert-base-swedish-cased", height = 800)
```

## SSYK and Swedish BERT examples:{visibility="uncounted"}

```{r}
library(tibble)
library(gt)

# New tibble with the provided data
new_occupation_data <- tibble(
    Rank = c(1, 2, 3, 4, 5),
    Closest_Match_Title_and_Code = c(
        "5326 - Ambulanssjukvårdare",
        "2226 - Ambulanssjuksköterskor m.fl.",
        "2227 - Geriatriksjuksköterskor",
        "5112 - Tågvärdar och ombordansvariga m.fl.",
        "2231 - Operationssjuksköterskor"
    )
)

# Create a gt table from the new tibble
new_occupation_table <- new_occupation_data %>%
    gt() %>%
    tab_header(
        title = md("**Closest SSYK Codes** to 'Ambulanssjukvårdare' (Code: 5326)")
    ) %>%
    cols_label(
        Rank = "Rank",
        Closest_Match_Title_and_Code = "Closest Match Title and Code"
    )

new_occupation_table %>% 
  tab_options(table.font.size = 40)

```

## SSYK and Swedish BERT examples:{visibility="uncounted"}

```{r}
library(tibble)
library(gt)

# New tibble with the provided data
new_occupation_data <- tibble(
    Rank = c(1, 2, 3, 4, 5),
    Closest_Match_Title_and_Code = c(
        "3115 - Ingenjörer och tekniker inom kemi och kemiteknik",
        "2145 - Civilingenjörsyrken inom kemi och kemiteknik",
        "3113 - Ingenjörer och tekniker inom elektroteknik",
        "7321 - Prepresstekniker",
        "2163 - Planeringsarkitekter m.fl."
    )
)

# Create a gt table from the new tibble
new_occupation_table <- new_occupation_data %>%
    gt() %>%
    tab_header(
        title = md("**Closest Matches** to 'Processingenjör, kemiteknik (Code: 2512)'")
    ) %>%
    cols_label(
        Rank = "Rank",
        Closest_Match_Title_and_Code = "Closest Match Title and Code"
    )

new_occupation_table %>% 
  tab_options(table.font.size = 40)
```

## Can we score how successful this is?{visibility="uncounted"}

We can use a crosswalk from one schema to another as a test set, e.g. O * Net to ISCO:

![](assets/scoring_assignement.png){fig-align="center" width=100%}